Information Closure Theory of Consciousness

### 2019-09-23 Paper_neural_coarse_graining_of_consciousness.tex
Updates from Overleaf

		%% our argument
		In this article, we propose a new information-based theory of consciousness, called \ac{OurTheory}. We argue that every process with a high degree of non-trivial information closure (NTIC) has consciousness. This means that the state of such a process corresponds one-to-one to conscious content.\footnote{In the following IC stands for "informational closure" or "informationally closed" and NTIC stands for "non-trivial informational closure" or "non-trivially informationally closed".}. We further postulate that the \textit{level} of consciousness corresponds to the degree of NTIC. (for the level versus content of consciousness see \cite{laureys2005neural, overgaard2010neural}).

@@

		\begin{figure}[H]
		    \centering
			\includegraphics[width=\textwidth]{WritingMaterials/Fig_ScaleProblemOfConsciousness/ScaleProblemOfConsciousness.pdf}
			\caption{The scale problem of consciousness: Human conscious experience does not reflect information processed at every scale. Only information at a certain coarse-grained scale in the neural system is conscious.}
			\label{fig:scaleproblem}
	   	\end{figure}

% ============================================================================ %
%                       Non-trivial informational closure                      %
% ============================================================================ %
	\section{Non-trivial informational closure} \label{sec:Non-trivial informational closure}
		The notion of non-trivial informational closure (NTIC) is introduced by \cite{BERTSCHINGER.2006}. The concept of closure is closely related to system identification in systems theory. One can distinguish a system from its environment by computing the closeness of the system \citep{maturana1991autopoiesis, rosen1991life, pattee2012evolving, luhmann1995probleme}. The closure can be further quantified by information theory.

@@

			\noindent
			This implies the system contains in itself all the information about its own future and the self-predictive information is gained from the information about the environment. Therefore, to form NTIC, the system can internalise and synchronise with the dynamics of the environment, i.e. modelling the environment. Furthermore, having high degrees of NTIC entails having high predictive power about the environment. This gives biological agents a great functional and evolutionary advantage.

			% Practically, \cite{guttenberg2016neural} proposed that NTIC can be formed by maximising the predictive power about the environment and the self-predictive information of the system concurrently.

			% Biological evidence of informational closure in the neural system
            % 			Despite the significant advantage for biological systems to form NTIC, to our knowledge, there are only a few studies directly examining NTIC in biological systems. Recent studies have shown relevant properties in the visual system of salamanders. Two studies \citep{Palmer2015, sederberg2018learning} examined salamander retina and found that the a large group of neural populations of retinal ganglion cells encoded predictive information about external stimuli also had high self-predictive information about their own future states (see Fig.~\ref{fig:sederberg2018learning}), therefore, in line with the characteristic of NTIC. Therefore, the self-predictive information of the neural populations can guide prediction about external stimuli without reference to stimulus parameters explicitly. Moreover, the internal predictability can be generalised to different stimulus classes suggesting the tendency to internalise the information of environmental dynamics in the neural populations. \toWrite{Add Some ending sentences}

            % 			\begin{figure}[H]
            % 				\includegraphics[width=\textwidth]{WritingMaterials/Fig_Sederberg/Fig_Sederberg.pdf}
            % 				\caption{Figure adapted from \cite{sederberg2018learning}}
            % 				\label{fig:sederberg2018learning}
            % 			\end{figure}

% ============================================================================ %
%                            Neural coarse-graining                            %
% ============================================================================ %
	\section{Coarse-graining in The Neural System} \label{sec:Neural coarse-graining}

		%% stochasticity at microscopic levels
		Despite evolutionary advantage of having NTIC processes in a system, however, the formation of NTIC witha  highly stochastic process is challenging. NTIC requires the predictability of the system state and is therefore impeded by noise in the system. Information processing at the microscopic levels (the cellular levels) in neural systems suffer from multiple environmental noise sources such as sensor, cellular, electrical, and synaptic noises. For example, neurons exhibit large trial-to-trial variability at the cellular level, and are subject to thermal fluctuations and other physical noises \citep{faisal2008noise}.

        %% Our argument
		However, we argue that, neural systems can form NTIC at certain macroscopic levels through coarse-graining of microscopic neural states. Coarse-graining refers to many-to-one functions which aggregate microscopic states to a macroscopic state. In other words, a number of different micro-states correspond to the same value of the macro-variable \citep{price2007causation}. Neural systems, therefore, can form more stable and deterministic state transitions and be more capable of forming NTIC processes at certain coarse-grained macroscopic levels.

		Indeed, empirical evidence suggests that coarse-graining is a common coding strategy to establish robustness against noise at microscopic levels of the the neural system. For instance, the inter-spike intervals of individual neurons are stochastic. This implies that the state of individual neurons cannot be used for encoding and representing stable information. However, the firing rate, i.e. the average spike counts over a given time interval, is more stable and robust against noise such as the variability in inter-spike intervals. Using this temporal coarse-graining strategy, known as the rate coding \citep{adrian1926impulses, gerstner2002spiking, maass2001pulsed, panzeri2015neural, stein2005neuronal}, neurons can encode stimulus intensity by increasing or decreasing the firing rate \citep{kandel2000principles}. \citep{stein2005neuronal}. The robustness of the rate coding is a direct consequence of the many to one mapping (i.e., coarse-graining).

		% population code
		Population coding is another example of encoding information through coarse-graining in neural systems. In this coding scheme, information is encoded by activation patterns of a set of neurons (a neuron population). In the population coding scheme, many states for a neuron population map to the same information content, thereby reducing the influence of noise in individual neurons. That is, stable representations can be formed through coarse-graining a high dimensional state space of a neuron population to a low dimensional state space of encoded contents \citep{kristan1997population, pouget2000information, binder2009encyclopedia, QuianQuiroga2009}. Therefore, individual neuron states (the microscopic level) are not informative about the encoded contents at the population level (the macroscopic level). Instead, coarse-grained variables are better substrates for stably encoding information and allow the neural system to ignore noisy interactions at the fine-grained level \citep{Woodward2007-WOOCWA}.

@@

        The first assumption is that every human conscious precept is informative. Here, "informative" refers to the resolution of uncertainty. Being in a certain conscious state rules out a number of other possible conscious states. This is in line with the impression that our conscious experience consist of multi-dimensional information and usually involves information from multiple modalities. Therefore, every human conscious precept resolves some amount of uncertainty and provides information.

        This assumption is also in agreement with one of the "axiom", \textit{information}, in Integrated Information Theory (IIT 3.0) which claims that \myQuote{...an experience of pure darkness is what it is by differing, in its particular way, from an immense number of other possible experiences...} \citep[p.2]{oizumi2014phenomenology}.

@@

        The third assumption is that no consciousness can contain information of the entire universe. This means that there exists a informational boundary for every consciousness, i.e. informationally closed to the the universe. This is in line with our conscious experience that information provided by every conscious percept is limited.

@@

        Note that a coarse-grained process $Y$ can be a coarse-graining of another coarse-grained process $S$ of the universe $X$ (see Fig.~\ref{fig:fullgraph}). Here, we consider $S$ is a dynamic system and $Y$ is a coarse-grained process of the system $S$. Therefore, the environment $E$ of $S$ is also a coarse-graining of the universe. Importantly, in such case, if a macroscopic process $Y$ is informationally closed to the universe $X$, this implies $Y$ is also informationally closed to the microscopic level of the system $S$ and the environment $E$.

@@

        This results in a crucial implication that informational closure can occur between levels in a system. The macroscopic level $Y$ can be informationally closed to the microscopic level $S$ within a system \citep{PFANTE.2014}.

@@

        A concrete example in the context of neuroscience is that $X$, $S$, $Y$, and $E$ represent the microscopic level of the universe, a cellular level process in the neural system, a macroscopic process of the neural system coarse-grained from the cellular level process $S$, and the environment which the cellular level process $S$ interacts with which may includes other processes in the neural system, the sensors for perception and interoception, and external physical worlds, respectively.

        Our fourth assumption is that information provided by every conscious content is not trivial. More precisely, there exists a certain degree of mutual information between conscious processes and the environments which it interacts with. As mentioned in Sec.\ref{sec:Non-trivial informational closure} regardless one process is informationally closed to another one, to form mutual information between the two processes, i.e. NTIC, is still possible. Therefore, a macroscopic process $Y$ can be NTIC to the system's environment $E$ and concurrently be informationally closed to its microscopic process $S$. In other words, the macroscopic process can encode environmental information (by forming NTIC) and, however, be ignorant to its microscopic processes (IC). This is in line with our conscious experience that information that every conscious percept provides represent rich and structured environmental states without involving much information at microscopic activities.

@@

		    %For example, a human brain may be a part of a conscious process at human society level. However, due to the informational closure, we should have no information about it in our consciousness.

@@

		=================================== Old  ===================================

	    There are nearly no have no commonality between GWT and \ac{OurTheory}. In GWT, only information can be accessed by GW is conscious. However, it's always difficult to identify the GW because there is no clear computational definition for it. In \ac{OurTheory}, any process forming NTIC is conscious. Therefore, to determine whether a process is conscious is a decidable problem for \ac{OurTheory} but not for GWT.

	    GWT claims that a module is not accessed by the GW is unconscious.

		Currently, we don't have strong theoretical implications to relate \ac{OurTheory} to GWT. However, based on \ac{OurTheory},  reasoning

		First, according to \ac{OurTheory}, consciousnesses occurs together with NTIC processes. Due to processes in the GW are accompanied by conscious experience, we suggest that the GW should occur at the scale where NTIC is formed. Second, notion of assess to GWS could be the state changes of the NTIC process at this coarse-grained level. Third, the all-or-none non-linear ignition may suggest the functions of coarse-graining are nonlinear. Finally, because NTIC encodes complex information about environmental states and dynamics,any state of the NTIC process can be seen as an integration of high dimensional information.

		P-consciousness vs A=consciousness

		Due to processes in the GW are accompanied by conscious experience, we can only postulate that the GW should occur at the scale where NTIC is formed. It is also likely that the

		GWT proposes a number of core properties, e.g., stability, broadcasting, and non-linear ignition.

		In \ac{OurTheory}, self-predictive information $I(Y_{t+1};Y_{t})$ is one of the quantity that contributes NTIC (Eq.~\ref{eq:nticObjective}). This property grantees the stability of a NTIC process. This outcome coincides with one of the GWT's claim that conscious contents often show stronger stability which may be necessary for the neural system to integrate information from a variety of modules. Although stability is claimed by both theories, the underliing reasons are different. GWT suggests that the neural mechanism to achieve stability is top-down amplifications from the GW to individual modules which gain the access to the GW. In contrast, \ac{OurTheory} suggests that stability is a natural consequence of NTIC without additional assumptions.

		We will address this comparison in our following modelling paper.
		Currently, we do not have concrete prediction about how an NTIC process reacts to rapid changes in the environment.
		More modeling work is needed

		However, regarding broadcasting and non-linear ignition

		conjectures to bridge the two theories.

		GW itself could be the NTIC process

		similar or dissimilar

		We suggests that the GW should occur at the scale where NTIC is formed. 
		We further hypothesise that the GW recruits the NTIC process to achieve its functional goal and, therefore, is accompany by conscious experience.

		Self-predictive information $I(Y_{t+1};Y_{t})$ is one of the quantity that contributes  NTIC (Eq.~\ref{eq:nticObjective}). This property grantees the stability of a NTIC process. This outcome coincides with one of the GWT's claim that conscious contents often show stronger stability which may be necessary for the neural system to integrate information from a variety of modules.Although stability is claimed by both theories, the proposed reasons are different. GWT suggests that the neural mechanism to achieve stability is top-down amplifications from the GW to individual modules which gain the access to the GW. In contrast, \ac{OurTheory} suggests that stability is a natural consequence of NTIC withour additional assumptions.

		The notion of assess to GWS could be the access at the coarse grained level.

		We suggest that when a coarse-grained variable becomes a part of the NTIC process. This makes information about the states of the variable occur in the conscious contents.

@@

% ============================================================================ %
%                                  Conclusion                                  %
% ============================================================================ %
    \section{Conclusion}
	\todo{@Ryota, sorry that I didn't notify you I will rewrite the conclusion entirely. This is the new version. Please have a look.}
    % short re-introduction 
    In this paper, we introduced \textbf{\acf{OurTheory}}, a new informational theory of consciousness. Based on four assumptions, \ac{OurTheory} proposes that a process which forms \textbf{non-trivial informational closure (NTIC)} is sufficient to be conscious and through coarse-graining the neural system forms a NTIC process, i.e., a conscious process, at a certain macroscopic level. \ac{OurTheory} considers that information is a common language to bridge the gap between conscious experience and the physical reality. Using information theory, \ac{OurTheory} proposes computational definitions for both conscious level and conscious content. This makes \ac{OurTheory} be able to generalise to any system beyond the human brains.

>>>>

		%% our argument
		In this article, we propose a new information-based theory of consciousness, called \ac{OurTheory}. We argue that every process with a high degree of non-trivial information closure (NTIC) has consciousness. This means that the state of such a process corresponds one-to-one to conscious content.\footnote{In the following IC stands for "informational closure" or "informationally closed" and NTIC stands for "non-trivial informational closure" or "non-trivially informationally closed".}. We further postulate that the \textit{level} of consciousness corresponds to the degree of NTIC. (for a discussion of the distinction between level versus content of consciousness see \cite{laureys2005neural, overgaard2010neural}).

@@

		\begin{figure}[H]
		    \centering
			\includegraphics[width=\textwidth]{WritingMaterials/Fig_ScaleProblemOfConsciousness/ScaleProblemOfConsciousness.pdf}
			\caption{The scale problem of consciousness: Human conscious experience does not reflect information from every scale. Only information at a certain coarse-grained scale in the neural system is reflected in consciousness.}
			\label{fig:scaleproblem}
	   	\end{figure}

% ============================================================================ %
%                       Non-trivial informational closure                      %
% ============================================================================ %
	\section{Non-trivial informational closure} \label{sec:Non-trivial informational closure}
		The notion of non-trivial informational closure (NTIC) is introduced by \cite{BERTSCHINGER.2006}. The concept of closure is closely related to system identification in systems theory. One can distinguish a system from its environment by computing the closedness of the system \citep{maturana1991autopoiesis, rosen1991life, pattee2012evolving, luhmann1995probleme}. The closure can be further quantified by information theory.

@@

			\noindent
			This implies the system contains in itself all the information about its own future and the self-predictive information contains the information about the environment. Therefore, to form NTIC, the system can internalise and synchronise with the dynamics of the environment, i.e. model the environment. Furthermore, having high degrees of NTIC entails having high predictive power about the environment. This gives biological agents a great functional and evolutionary advantage.

% ============================================================================ %
%                            Neural coarse-graining                            %
% ============================================================================ %
	\section{Coarse-graining in the Neural System} \label{sec:Neural coarse-graining}

		%% stochasticity at microscopic levels
		The formation of NTIC with a  highly stochastic process is challenging. NTIC requires the predictability of the system state and is therefore impeded by noise in the system. Information processing at the microscopic levels (the cellular levels) in neural systems suffer from multiple environmental noise sources such as sensor, cellular, electrical, and synaptic noises. For example, neurons exhibit large trial-to-trial variability at the cellular level, and are subject to thermal fluctuations and other physical noises \citep{faisal2008noise}.

        %% Our argument
		However, it is possible that neural systems form NTIC at certain macroscopic levels through coarse-graining of microscopic neural states.

		Coarse-graining refers to many-to-one functions which aggregate microscopic states to a macroscopic state. In other words, a number of different micro-states correspond to the same value of the macro-variable \citep{price2007causation}. Coarse-grainings, can therefore form more stable and deterministic state transitions and more often form NTIC processes.

		For neural systems this means that a microscopically noisy neural system may still give rise to an NTIC process
		on a more macroscopic scale.

		In neural systems such coarse-grainings have been observed to represent external stimuli in empirical studies.

		=== so, to here, it's fine ===
		Because coarse-graining is an effective way to create stable variables, a stochastic system (e.g., neural systems) can represent information stably by encoding information using coarse-grained macroscopic variables.

		============================================================================

		Indeed, empirical evidence suggests that coarse-graining is a common coding strategy to establish robustness against noise at microscopic levels of the neural system. For instance, the inter-spike intervals of an individual neuron are stochastic. This implies that the state of an individual neuron does not represent stable information. However, the firing rate, i.e. the average spike counts over a given time interval, is more stable and robust against noise such as the variability in inter-spike intervals. Using this temporal coarse-graining strategy, known as rate coding \citep{adrian1926impulses, gerstner2002spiking, maass2001pulsed, panzeri2015neural, stein2005neuronal}, neurons can encode stimulus intensity by increasing or decreasing the firing rate \citep{kandel2000principles}. \citep{stein2005neuronal}. The robustness of the rate coding is a direct consequence of the many to one mapping (i.e., coarse-graining).

		% population code
		Population coding is another example of encoding information through coarse-graining in neural systems. In this coding scheme, information is encoded by activation patterns of a set of neurons (a neuron population). In the population coding scheme, many states for a neuron population map to the same state of encoded information, thereby reducing the influence of noise in individual neurons. That is, stable representations can be formed through coarse-graining the high dimensional state space of a neuron population to a lower dimensional macroscopic state space \citep{kristan1997population, pouget2000information, binder2009encyclopedia, QuianQuiroga2009}. Therefore, individual neuron states (the microscopic level) are not informative enough about the full encoded contents at the population level (the macroscopic level). Instead, coarse-grained variables are better substrates for stably encoding information and allow the neural system to ignore noisy interactions at the fine-grained level \citep{Woodward2007-WOOCWA}.

@@

        The first assumption is that every human conscious percept is informative. Here, "informative" refers to the resolution of uncertainty. Being in a certain conscious state rules out a number of other possible conscious states. This is in line with the impression that our conscious experience consist of multi-dimensional information and usually involves information from multiple modalities. Therefore, every human conscious percept resolves some amount of uncertainty and provides information.

        This assumption is also in agreement with the "axiom" of \textit{information} in Integrated Information Theory (IIT 3.0) which claims that \myQuote{...an experience of pure darkness is what it is by differing, in its particular way, from an immense number of other possible experiences...} \citep[p.2]{oizumi2014phenomenology}.

@@

        === How about if I change this ===
        The third assumption is that no consciousness can contain complete information about the universe. This means that there exists a informational boundary for every consciousness, i.e. informationally closed to the the universe. This is in line with our conscious experience that information provided by every conscious percept is limited.

        === to this ===
        The third assumption is that there exists a informational boundary for every consciousness. The informational boundary here means that no consciousness can contain information of the entire universe and every consciousness is informationally closed to the the universe. This is in line with our conscious experience that information provided by every conscious percept is limited.

@@

        Note that a coarse-grained process $Y$ can be a coarse-graining of another coarse-grained process $S$ of the universe $X$ (see Fig.~\ref{fig:fullgraph}). Here, we consider $S$ as a coarse-graining of $X$ and call it the system. Then $Y$ is a coarse-graining of the system $S$. We can then also define another coarse-graining of $X$ called the environment $E$ of $S$. This is the coarse-graining that contains the information about $X$ that is not contained in $S$. Importantly, in this case, if a macroscopic process $Y$ is informationally closed to the universe $X$, this implies $Y$ is also informationally closed to both the system $S$ and the environment $E$.

@@

        This results in a crucial implication that informational closure can occur between levels in a system. The macroscopic level $Y$ can be informationally closed to its microscopic level $S$ within a system \citep{PFANTE.2014}.

@@

        A concrete example in the context of neuroscience is that $X$ represents the microscopic level of the universe, $S$ a cellular level process in the neural system, $Y$ a more macroscopic process of the neural system coarse-grained from the cellular level process $S$, and $E$ the environment which the cellular level process $S$ interacts with. The environment $E$ may includes other processes in the neural system, the sensors for perception and interoception, and external physical worlds, respectively.

        Our fourth assumption is that information provided by every conscious content is not trivial. More precisely, there exists a certain degree of mutual information between conscious processes and the environments which it interacts with. As mentioned in Sec.\ref{sec:Non-trivial informational closure} a process can be informationally closed to another one and still have mutual information with it, i.e. be NTIC. Therefore, a macroscopic process $Y$ can be NTIC to the system's environment $E$ and at the same time be informationally closed to its microscopic process $S$. In other words, the macroscopic process can encode environmental information (by forming NTIC) and, however, be ignorant to its microscopic processes (IC). This is in line with our conscious experience that information that every conscious percept provides represent rich and structured environmental states without involving much information at microscopic activities.

@@

@@

% 		=================================== Old  ===================================

% 	    There are nearly no have no commonality between GWT and \ac{OurTheory}. In GWT, only information can be accessed by GW is conscious. However, it's always difficult to identify the GW because there is no clear computational definition for it. In \ac{OurTheory}, any process forming NTIC is conscious. Therefore, to determine whether a process is conscious is a decidable problem for \ac{OurTheory} but not for GWT.

% 	    GWT claims that a module is not accessed by the GW is unconscious.

% 		Currently, we don't have strong theoretical implications to relate \ac{OurTheory} to GWT. However, based on \ac{OurTheory},  reasoning

% 		First, according to \ac{OurTheory}, consciousnesses occurs together with NTIC processes. Due to processes in the GW are accompanied by conscious experience, we suggest that the GW should occur at the scale where NTIC is formed. Second, notion of assess to GWS could be the state changes of the NTIC process at this coarse-grained level. Third, the all-or-none non-linear ignition may suggest the functions of coarse-graining are nonlinear. Finally, because NTIC encodes complex information about environmental states and dynamics,any state of the NTIC process can be seen as an integration of high dimensional information.

% 		P-consciousness vs A=consciousness

% 		Due to processes in the GW are accompanied by conscious experience, we can only postulate that the GW should occur at the scale where NTIC is formed. It is also likely that the

% 		GWT proposes a number of core properties, e.g., stability, broadcasting, and non-linear ignition.

% 		In \ac{OurTheory}, self-predictive information $I(Y_{t+1};Y_{t})$ is one of the quantity that contributes NTIC (Eq.~\ref{eq:nticObjective}). This property grantees the stability of a NTIC process. This outcome coincides with one of the GWT's claim that conscious contents often show stronger stability which may be necessary for the neural system to integrate information from a variety of modules. Although stability is claimed by both theories, the underliing reasons are different. GWT suggests that the neural mechanism to achieve stability is top-down amplifications from the GW to individual modules which gain the access to the GW. In contrast, \ac{OurTheory} suggests that stability is a natural consequence of NTIC without additional assumptions.

% 		We will address this comparison in our following modelling paper.
% 		Currently, we do not have concrete prediction about how an NTIC process reacts to rapid changes in the environment.
% 		More modeling work is needed

% 		However, regarding broadcasting and non-linear ignition

% 		conjectures to bridge the two theories.

% 		GW itself could be the NTIC process

% 		similar or dissimilar

% 		We suggests that the GW should occur at the scale where NTIC is formed. 
% 		We further hypothesise that the GW recruits the NTIC process to achieve its functional goal and, therefore, is accompany by conscious experience.

% 		Self-predictive information $I(Y_{t+1};Y_{t})$ is one of the quantity that contributes  NTIC (Eq.~\ref{eq:nticObjective}). This property grantees the stability of a NTIC process. This outcome coincides with one of the GWT's claim that conscious contents often show stronger stability which may be necessary for the neural system to integrate information from a variety of modules.Although stability is claimed by both theories, the proposed reasons are different. GWT suggests that the neural mechanism to achieve stability is top-down amplifications from the GW to individual modules which gain the access to the GW. In contrast, \ac{OurTheory} suggests that stability is a natural consequence of NTIC withour additional assumptions.

% 		The notion of assess to GWS could be the access at the coarse grained level.

% 		We suggest that when a coarse-grained variable becomes a part of the NTIC process. This makes information about the states of the variable occur in the conscious contents.

		% ================================= Super old ================================ %

@@

% ============================================================================ %
%                                  Conclusion                                  %
% ============================================================================ %
    \section{Conclusion}
    % short re-introduction 
    In this paper, we introduced \textbf{\acf{OurTheory}}, a new informational theory of consciousness. Based on four assumptions, \ac{OurTheory} proposes that a process which forms \textbf{non-trivial informational closure (NTIC)} is sufficient to be conscious and through coarse-graining the neural system forms a NTIC process, i.e., a conscious process, at a certain macroscopic level. \ac{OurTheory} considers that information is a common language to bridge the gap between conscious experience and the physical reality. Using information theory, \ac{OurTheory} proposes computational definitions for both conscious level and conscious content. This makes \ac{OurTheory} be able to generalise to any system beyond the human brains.
